---
title: "Polynomial , KNN and Regression Models"
author: "Prabin Kharel"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Supervised Learning with Polynomial, KNN and ANN regression model

##### Loading covid_tbl_final.xslx

```{r}
library(readxl)
covid_data <- read_xlsx("/home/prabin/Downloads/covid_tbl_final.xlsx")
head(covid_data)
```

##### Cleaning covid_data

```{r}
str(covid_data)
```

```{r}
covid_data$Date<-as.Date(as.POSIXct(covid_data$Date))
str(covid_data)
```

##### Plotting total deaths vs date

```{r}
plot(covid_data$Date,covid_data$Deaths_total, xlab="Date",ylab="Total Deaths")
```

##### Plotting daily deaths vs date

```{r}
plot(covid_data$Date,covid_data$Deaths_daily,main="Daily Deaths from 2020-01-23 to 2020-05-31", xlab="Date", ylab = "Daily Deaths")
```

```{r}
summary(covid_data$Deaths_daily)
```

```{r}
library(dplyr)
filter(covid_data, Deaths_daily >=50 & Date <=as.Date("2021-03-05"))

```

```{r}
wsn<-c(399,401,408)
for(i in 1:length(wsn)){

temp_sn = wsn[i]  
# Get the Value to be adjusted
curr_val<-covid_data[covid_data$SN==temp_sn,"Deaths_daily"]
# Calculate the average daily deaths for last 30 days
avg_daily_deaths<-ceiling(mean(covid_data[covid_data$SN %in% c((temp_sn-1):(temp_sn-1-30)),]$Deaths_daily))

# Change the Value for given SN
covid_data[covid_data$SN==temp_sn,"Deaths_daily"]=avg_daily_deaths
# Change values for last 30 days
covid_data[covid_data$SN %in% c((temp_sn-1):(temp_sn-1-30)),]$Deaths_daily=as.integer( round(curr_val/30))
}
```

```{r}
plot(covid_data$Date,
covid_data$Deaths_daily,
main = "Daily Deaths: 23 Jan 2020
- 31 May 2021",
xlab = "Date",
ylab = "Daily Deaths")
```

      Since the data is now clean, we divide the data into training and testing set for our regression model.
      
##### Splitting the data into training and testing set
```{r}
set.seed(1234)
ind <- sample(2, nrow(covid_data), replace=T, prob=c(0.7,0.3))
train_data <- covid_data[ind==1,]
test_data <- covid_data[ind==2,]
```


##### Linear regression model

```{r}
library(caret)
lm1 <- train(Deaths_daily~SN, data = train_data, method="lm")
summary(lm1)
predict1 <- predict(lm1, newdata = test_data)

```

######Evaluating metrics
```{r}
R2 <- R2(predict1,test_data$Deaths_daily)
RMSE <- RMSE(predict1,test_data$Deaths_daily)
MAE <- MAE(predict1,test_data$Deaths_daily)
R2
RMSE
MAE

```

##### Quadratic linear regression model

```{r}
lm2 <- train(Deaths_daily~poly(SN,2),data=train_data, method="lm")
summary(lm2)
predict2 <- predict(lm2,newdata = test_data)

```
```{r}
head(predict2)
```

```{r}
R2 <- R2(predict2,test_data$Deaths_daily)
RMSE <- RMSE(predict2,test_data$Deaths_daily)
MAE <- MAE(predict2,test_data$Deaths_daily)
R2
RMSE
MAE
```

##### Cubic linear regression model
```{r}
lm3 <- train(Deaths_daily~poly(SN,3), data = train_data, method="lm")
summary(lm3)
```

```{r}
predict3 <- predict(lm3,newdata = test_data)
head(predict3)
```
```{r}
R2 <- R2(predict3,test_data$Deaths_daily)
RMSE <- RMSE(predict3,test_data$Deaths_daily)
MAE <- MAE(predict3,test_data$Deaths_daily)
R2
RMSE
MAE
```


##### Double quadratic linear model

```{r}
lm4 <- train(Deaths_daily~poly(SN,4), data=train_data, method="lm")
summary(lm4)
```

```{r}
predict4 <- predict(lm4, newdata = test_data)
head(predict4)
```
```{r}
R2 <- R2(predict4,test_data$Deaths_daily)
RMSE <- RMSE(predict4,test_data$Deaths_daily)
MAE <- MAE(predict4,test_data$Deaths_daily)
R2
RMSE
MAE
```
##### Fifth order polynomial regression model
```{r}
lm5 <- train(Deaths_daily ~SN, data= train_data, methos="lm")
summary(lm5)
```

```{r}
predict5 <- predict(lm5, newdata = test_data)
head(predict5)
```

```{r}
R2 <- R2(predict5,test_data$Deaths_daily)
RMSE <- RMSE(predict5,test_data$Deaths_daily)
MAE <- MAE(predict5,test_data$Deaths_daily)
R2
RMSE
MAE
```





##### KNN regression model
```{r}
knnmodel <- train(Deaths_daily ~SN, data=train_data, method="knn")
summary(knnmodel)
```

```{r}
predict6 <- predict(knnmodel, newdata = test_data)
head(predict6)
```
```{r}
R2 <- R2(predict6,test_data$Deaths_daily)
RMSE <- RMSE(predict6,test_data$Deaths_daily)
MAE <- MAE(predict6,test_data$Deaths_daily)
R2
RMSE
MAE
```

##### ANN-MLP regression model with 2 hidden layers with 3 and 2 neurons

```{r message=FALSE}
#install.packages("neuralnet")
library(neuralnet)
neural1 <- neuralnet(Deaths_daily ~ SN, data= train_data, hidden=c(3,2),linear.output = F)
plot(neural1, main="Neural network with 2 hidden layers with 3 and 2 neurons")
summary(neural1)
```

```{r}
predict7 <- predict(neural1, newdata = test_data)
head(predict7)
```
```{r}
R2 <- R2(predict7,test_data$Deaths_daily)
RMSE <- RMSE(predict7,test_data$Deaths_daily)
MAE <- MAE(predict7,test_data$Deaths_daily)
R2
RMSE
MAE
```
##### Selection of best model based on RMSE on test data

Comparing the RMSE of all the models above , we see that RMSE of the KNN model is the least. So, KNN model is our best model.

##### Summary and Recommendation

```{r}
plot(covid_data$SN, covid_data$Deaths_daily,
main = "Daily Covid Deaths",
xlab = "SN",
ylab = "Daily Deaths")
lines(predict(knnmodel,newdata = covid_data), col = "red", lwd=2)
```

From the prediction from our best model and the data in hand it is safe to say that there will be a rise in the cases for a certain period. What could be advised on this is that the ministry of health provide vaccination and maintain awareness among the people. Since more deaths would mean more infected people and more carriers(although they might not show symptoms or be medically diagnosed) the government should find a way to impose lockdown and minimal crowd gatherings keeping in mind a way to fulfill the basic requirements of the people for the time being.